Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller{Petersburg version \YellowPaperVersionNumber}

### 2021-04-11 Paper.tex
Update `SSTORE` cost and refund for EIP-2200
This patch ports the decision tree from the specification of [EIP-2200]
into the instruction set reference. Fixes #788.

[EIP-2200]: https://eips.ethereum.org/EIPS/eip-2200

wchargin-branch: eip-2200
wchargin-source: 92c2af76d68bd0da255770fa4578091342835ef0

\begin{tabu}{l r l}
\toprule
Name & Value & Description* \\
\midrule
$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
$G_{extcodehash}$ & 400 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
&&contract. \\

@@

&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
G_{sreset} & \text{otherwise}

@@

R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\

>>>>

\begin{tabu}{l r l}
\toprule
Name & Value & Description* \\
\midrule
$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
$G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
$G_{extcodehash}$ & 700 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
$G_{balance}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
$G_{sload}$ & 800 & Paid for a {\small SLOAD} operation. \\
$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
&&is set to zero. \\
$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
&&non-zero.\\
\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
&&contract. \\

@@

&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{r}$ are specified by EIP-2200 as follows. \\
&&&& Let $\boldsymbol{\sigma}^0$ be the state if the current transaction were to revert (the ``original state''). \\
&&&& Let $v^0 = \boldsymbol{\sigma}^0[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the original value of the storage slot. \\
&&&& Let $v = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the current value. \\
&&&& Let $v' = \boldsymbol{\mu}_{\mathbf{s}}[1]$ be the new value. \\
&&&& Then: \\
&&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
G_{sload} & \text{if} \quad v = v' \\
G_{sset}  & \text{if} \quad v^0 = v \; \wedge \; v^0 = 0 \\
G_{sreset}  & \text{if} \quad v^0 = v \; \wedge \; v^0 \neq 0 \\
G_{sload} & \text{otherwise}

@@

0 & \text{if} \quad v = v' \\
0 & \text{if} \quad v^0 = v \; \wedge \; v^0 = 0 \\
R_{sclear} & \text{if} \quad v^0 = v \; \wedge \; v' = 0 \\
r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{otherwise}
&&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
-R_{sclear} & \text{if} \quad v^0 \neq 0 \; \wedge \; v = 0 \\
R_{sclear} & \text{if} \quad v^0 \neq 0 \; \wedge \; v' = 0 \\
0 & \text{otherwise} \\
&&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
G_{sset} - G_{sload}   & \text{if} \quad v^0 = v' \; \wedge \; v^0 = 0 \\
G_{sreset} - G_{sload} & \text{if} \quad v^0 = v' \; \wedge \; v^0 \neq 0 \\
0 & \text{otherwise} \\
